Resource Buying Games

### 2017-02-01 slides.tex
added explaining picture to construct game without pne
work work work
added theorem and explanation on non-decreasing marginal cost functions, decided to leave anarchy out since the explanation is rather dubious

\documentclass{beamer}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{cancel}
\usetheme{metropolis}
\title{Resource Buying Games}
\author[C. Welzel]{Christoph Welzel}

@@

\begin{document}
\input{tikz/settings}
\maketitle
\section{Introduction}
\begin{frame}
\end{frame}

\subsection{Game definition}
\begin{frame}
  \frametitle{Definition: congestion model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $N$: set of players
    \item $E$: set of resources
    \item $\mathcal{S} = \times_{i\in N}\mathcal{S}_{i}$ with $\mathcal{S}_{i}$
      set of desired resources $S_{i}\subseteq E$: set of configurations
    \item $d_{i}$: demand of player $i$ for her resource
      \begin{itemize}
        \item $d_{i} = 1$ for an unweighted game
        \item $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$: load of $e$
          in configuration $S\in\mathcal{S}$
      \end{itemize}
    \item $c_{e}:\mathbb{N}\rightarrow\mathbb{N}$ with
      $c_{e}(S) = c_{e}(\ell_{e}(S))$: cost for the resource $e$ under the load
      of $S$
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Operations on Matroids $\mathcal{M} = \tupel{E, \mathcal{B}}$}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{center}
      Contraction:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}/e = \tupel{E\setminus\set{e}, \mathcal{B}/e}\text{ with}\\
      &\mathcal{B}/e = \set{B\subseteq (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}
    \end{align*}
    \column<4->{0.5\textwidth}
    \begin{center}
      Deletion:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}\setminus e = \tupel{E\setminus\set{e},\mathcal{B}\setminus e}\text{ with}\\
      &\mathcal{B}\setminus e = \set{B\subseteq E\setminus\set{e}\middle|B\in\mathcal{B}}
    \end{align*}
  \end{columns}
  \begin{center}
    \begin{tikzpicture}
      \uncover<2->{\node (G) {\input{tikz/graphp1}};
      \node[right=1 of G] (leadsto) {\Huge$\leadsto$};}
      \uncover<3>{\node[right=1 of leadsto] (G') {\input{tikz/graphp1Cont}};}
      \uncover<5>{\node[right=1 of leadsto] (G'') {\input{tikz/graphp1Del}};}
    \end{tikzpicture}
  \end{center}
\end{frame}

@@

\begin{frame}
  \frametitle{Existence of Pure Nash Equilibria}
  \begin{theorem}
    Every weighted matroid resource buying game with marginally non-increasing
    cost functions possesses a pure Nash equilibria.
  \end{theorem}
  \begin{itemize}
    \item socially optimal configuration profile: $B\in\mathcal{B}$ such that
      $\sum_{e\in E}c_{e}(\ell_{e}(B))$ is minimal
    \item for socially optimal configuration profile $B$ holds:
      \begin{equation*}
        c_{e}(B)\leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
        \text{ for all }e\in E\setminus F
      \end{equation*}
    \item $B$ is a pure Nash equilibrium
    \item (but finding socially optimal configuration profiles is NP-hard)
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Construction of games without PNE}

\end{frame}

>>>>

\documentclass{beamer}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{cancel}
\usepackage{stmaryrd}
\usetheme{metropolis}
\title{Resource Buying Games}
\author[C. Welzel]{Christoph Welzel}

@@

\begin{document}
\input{tikz/settings}
\maketitle

\begin{frame}
  \frametitle{Introduction}
\end{frame}

\section{Basic Definitions}
\begin{frame}
  \frametitle{Definition: congestion model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $N$: set of players
    \item $E$: set of resources
    \item $\mathcal{S} = \times_{i\in N}\mathcal{S}_{i}$ with $\mathcal{S}_{i}$
      set of desired resources $S_{i}\subseteq E$: set of configurations
    \item $d_{i}$: demand of player $i$ for her resource
      \begin{itemize}
        \item $d_{i} = 1$ for an unweighted game
        \item $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$: load of $e$
          in configuration $S\in\mathcal{S}$
      \end{itemize}
    \item $c_{e}:\mathbb{N}\rightarrow\mathbb{N}$ with
      $c_{e}(S) = c_{e}(\ell_{e}(S))$: cost for the resource $e$ under the load
      of $S$
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Operations on Matroids $\mathcal{M} = \tupel{E, \mathcal{B}}$}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{center}
      Contraction:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}/e = \tupel{E\setminus\set{e}, \mathcal{B}/e}\text{ with}\\
      &\mathcal{B}/e = \set{B\subseteq (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}
    \end{align*}
    \column<4->{0.5\textwidth}
    \begin{center}
      Deletion:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}\setminus e = \tupel{E\setminus\set{e},\mathcal{B}\setminus \set{e}}\text{ with}\\
      &\mathcal{B}\setminus e = \set{B\subseteq E\setminus\set{e}\middle|B\in\mathcal{B}}
    \end{align*}
  \end{columns}
  \begin{center}
    \begin{tikzpicture}
      \uncover<2->{\node (G) {\input{tikz/graphp1}};
      \node[right=1 of G] (leadsto) {\Huge$\leadsto$};}
      \uncover<3>{\node[right=1 of leadsto] (G') {\input{tikz/graphp1Cont}};}
      \uncover<5>{\node[right=1 of leadsto] (G'') {\input{tikz/graphp1Del}};}
    \end{tikzpicture}
  \end{center}
\end{frame}

@@

\begin{frame}
  \frametitle{Existence of Pure Nash Equilibria}
  \begin{theorem}
    Every weighted matroid resource buying game with marginally non-increasing
    cost functions possesses a PNE.
  \end{theorem}
  \begin{itemize}
    \item socially optimal configuration profile: $B\in\mathcal{B}$ such that
      $\sum_{e\in E}c_{e}(\ell_{e}(B))$ is minimal
    \item for socially optimal configuration profile $B$ holds:
      \begin{equation*}
        c_{e}(B)\leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
        \text{ for all }e\in E\setminus F
      \end{equation*}
    \item for $B$ exists $p$ s.t. $\tupel{B,p}$ is a PNE
    \item (but finding socially optimal configuration profiles is NP-hard)
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Construction of games without PNE}
   \begin{columns}
     \column{0.3\textwidth}
       \begin{itemize}
         \item $c_{x}(t) = t$
         \item $c_{y}(t) = 5\frac{1}{2}$
         \item $c_{z}(t) = 4$
         \item $d_{1} = 5$
         \item $d_{2} = 4$
       \end{itemize}
     \column{0.7\textwidth}
       \begin{center}
         \resizebox{\textwidth}{!}{\input{tikz/gamediagram}}
       \end{center}
   \end{columns}
   \begin{itemize}
     \item assume PNE $\tupel{Z^{\ast} = \tupel{Z_{1}^{\ast}, Z_{2}^{\ast}},
       p^{\ast} = \tupel{p_{1}^{\ast}, p_{2}^{\ast}}}$
     \item $Z_{i}^{\ast}\subseteq\tupel{X_{i}\cup Y_{i}}$
     \item $x\not\in Z_{1}^{\ast}\Rightarrow \set{y,z}\subseteq Z_{1}^{\ast}$
     \item $y\not\in Z_{2}^{\ast}\Rightarrow \set{x,z}\subseteq Z_{2}^{\ast}$
     \item[$\Rightarrow$] neiter $x\in Z_{1}^{\ast}$ nor $x\not\in Z_{1}^{\ast}$ $\lightning$
   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Non-Decreasing Marginal Cost Functions}
  \begin{columns}
    \column{0.5\textwidth}
      Non-Increasing Marginal Cost Functions:
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{ln(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \column{0.5\textwidth}
      Non-Decreasing Marginal Cost Functions:
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              xmin=1,
              xmax=4,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{5*exp(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
  \end{columns}
  \begin{theorem}
    Let $G$ be a weighted resource buying game with non-decreasing marginal
    costs and let $S^{\ast}$ be a socially optimal solution. We can compute
    (so called) marginal cost pricing vectors which induce a PNE.
  \end{theorem}
\end{frame}
